Generalized Tractability

### 2023-06-05 GeneralizedTractability_230529.tex
Updates from Overleaf

A problem is defined as $T$-tractable with parameter $\vp$ iff there exists a positive constant $C_{\vp}$, which is independent of $\varepsilon$ and $d$, such that
\begin{equation} \label{eq:tract_def}
	\comp(\varepsilon,d) \le C_{\vp}\, T(\varepsilon^{-1},d,\vp) \qquad \forall \varepsilon >0, \ d \in \mathbb{N}.
\end{equation}
A problem is \emph{strongly}
$T$-tractable with parameter $\vp$ iff the complexity is independent of the dimension of the problem, that is, there exists a positive constant $C_{\vp}$, again independent of $\varepsilon$ and $d$, such that
\begin{equation} \label{eq:strong_tract_def}
	\comp(\varepsilon,d) \le C_{\vp}\, T(\varepsilon^{-1},1,\vp) \qquad \forall \varepsilon >0 , \ d \in \mathbb{N}.
\end{equation}
In this article we allow $\vp$ to be a scalar or vector as the situation dictates.

@@

If \eqref{eq:tract_def} holds for some $\vp$, it clearly holds for larger $\vp$.  We are often interested in the optimal or smallest $\vp$ for which \eqref{eq:tract_def} holds.  Define the closures of the sets of parameters for which our (strong) tractability conditions hold:
\begin{equation*}
	\cp_{\textup{trct}} : = \{\vp^* : \eqref{eq:tract_def} \text{ holds }\forall \vp \in (\vp^*,\boldsymbol{\infty})\}, \qquad
	\cp_{\textup{strct}} : = \{\vp^* : \eqref{eq:strong_tract_def} \text{ holds }\forall \vp \in (\vp^*,\boldsymbol{\infty})\}.
\end{equation*}
The  set of optimal parameters is defined as all of those parameters satisfying the (strong) tractability conditions that are not greater than or equal to others:
\begin{gather}
	\label{eq:Poptdef}
	\mathcal{P}_{\textup{opt}} : = \{\vp^* \in \cp_{\textup{trct}} :  \vp^* \notin [\vp,\boldsymbol{\infty}) \ \forall \vp \in  \cp_{\textup{trct}} \setminus \{\vp^*\} \}, \\
	\label{eq:Psoptdef}
	\mathcal{P}_{\textup{sopt}} : = \{\vp^* \in \cp_{\textup{strct}} :  \vp^* \notin [\vp,\boldsymbol{\infty}) \ \forall \vp \in  \cp_{\textup{strct}} \setminus \{\vp^*\} \}.
\end{gather}

@@

Suppose that the domain of interest $(\varepsilon^{-1},d)$ is not all of $(0,\infty) \times \naturals$ but some subset
\begin{equation}
    \Omega  := \{(\varepsilon^{-1},d) : \varepsilon > E(d) \ge 0, \ d \in \naturals \}, \qquad E : \naturals \to [0,\infty).
\end{equation}
We then expect that the equivalent condition for (strong) tractability will be similar in form, but weaker than, the conditions in Theorems  \ref{thm_main_strong_tract2} and \ref{thm_main_tract2}.
A problem is \peter{strongly}
$T$-tractable with parameter $\vp$ \emph{on the restricted domain $\Omega$} iff the complexity is independent of the dimension of the problem, that is, there exists a positive constant $C_{\vp}$, again independent of $\varepsilon$ and $d$, such that
\begin{equation} \label{eq:rest_strong_tract_def}
	\comp(\varepsilon,d) \le C_{\vp}\, T(\varepsilon^{-1},1,\vp) \qquad \forall(\varepsilon^{-1},d) \in \Omega.
\end{equation}
%\peter{[I am wondering whether this makes sense if $E$ depends on $d$? Do Gnewuch and Wo\'{z}niakowski allow dependence of $E$ on $d$?]}
A problem is 
$T$-tractable with parameter $\vp$ \emph{on the restricted domain $\Omega$} iff the complexity is independent of the dimension of the problem, that is, there exists a positive constant $C_{\vp}$, again independent of $\varepsilon$ and $d$, such that
\begin{equation} \label{eq:rest_tract_def}
	\comp(\varepsilon,d) \le C_{\vp}\, T(\varepsilon^{-1},d,\vp) \qquad \forall(\varepsilon^{-1},d) \in \Omega.
\end{equation}

Define 
\begin{equation} \label{eq:Mddef}
    M(d) := \begin{cases}
        \max\{n : \lambda_{n,d} > E(d)\}, & E(d) > 0, \\
        \infty, & E(d) = 0.
    \end{cases}
\end{equation}
This means that 
\[
\begin{cases}
\lambda_{1,d}\ge  \cdots \ge \lambda_{M(d),d} > E(d) \ge\lambda_{M(d)+1,d}, \quad
(\lambda_{1,d}^{-1},d), \ldots, (\lambda^{-1}_{M(d),d},d)  \in \Omega, & E(d) > 0, \\
\lambda_{1,d}\ge \lambda_{2,d}\ge \cdots  > 0, \quad
(\lambda_{1,d}^{-1},d), (\lambda_{2,d}^{-1},d), \ldots \in \Omega, & E(d) = 0.
\end{cases}
\]
Also, for each $d \in \naturals$ define the interval 
\begin{equation} \label{eq:hIdef}
\hI_{d}:=\begin{cases} 
\{1,\ldots, M(d)\}, & M(d) < \infty, \\
\naturals, & M(d) = \infty.
\end{cases}
\end{equation}
From the definition of $M(d)$ it follows that 
\[
\comp(\varepsilon,d) \in \hI_d \qquad \forall (\varepsilon^{-1},d) \in \Omega.
\]

\begin{theorem}\label{thm_main_rest_strong_tract}
Let $T$ be a tractability function as specified in \eqref{eq:Tspec} and satisfying \eqref{eq:Tconditions} and \eqref{eq:ptauassume}.  A problem is strongly $T$-tractable \emph{on the restricted domain $\Omega$} iff there exists $\vp \in [\vzero, \boldsymbol{\infty})$ and an integer $L_{\vp} > 0$ such that
\begin{equation} \label{eq:strong_rest_tractiff}
     S_{\vp}:=\sup_{d \in \naturals} \sum_{i = L_{\vp}}^{M(d)} \frac{1}{T(\lambda_{i,d}^{-1},1,\vp)} < \infty.
\end{equation}
If \eqref{eq:strong_rest_tractiff} holds for some $\vp$, let  $\widetilde{\cp}_{\textup{rstrct}} : = \{\vp^* : \eqref{eq:strong_rest_tractiff} \text{ holds }\forall \vp \in (\vp^*,\boldsymbol{\infty})\}$.  Then $\cp_{\textup{rstrct}} = \widetilde{\cp}_{\textup{rstrct}}$, and the set of optimal strong tractability parameters on the restricted domain is
\[
	\cp_{\textup{rsopt}} =
	\{\vp^* \in \widetilde{\cp}_{\textup{rstrct}} :  \vp^* \notin [\widetilde{\vp},\boldsymbol{\infty}) \ \forall \widetilde{\vp}\in  \widetilde{\cp}_{\textup{rstrct}} \setminus \{\vp^*\} \}.
\]
\end{theorem}

@@

\begin{theorem}\label{thm_main_rest_tract2}
Let $T$ be a tractability function as specified in \eqref{eq:Tspec} and satisfying \eqref{eq:Tconditions} and \eqref{eq:ptauassume}.  A problem is $T$-tractable on $\Omega$ iff there exists $\vp\ge \vzero$ and a positive integer $L_{\vp}$ such that
\begin{equation} \label{eq:rest_tractiff4}
     S_{\vp}:=\sup_{d \in \naturals}
     \sum_{i = \lceil L_{\vp}\, T(0,d,\vp) \rceil}^{M(d)} \frac{1}{T(\lambda^{-1}_{i,d},d,\vp)}< \infty.
\end{equation}
If \eqref{eq:rest_tractiff4} holds for some $\vp$, let $\widetilde{\cp}_{\textup{trct}} : = \{\vp^* : \eqref{eq:tractiff4} \text{ holds }\forall \vp \in (\vp^*,\boldsymbol{\infty}) \}$.
Then $\cp_{\textup{trct}} = \widetilde{\cp}_{\textup{trct}}$, and the set of optimal  tractability parameters is
\[
\cp_{\textup{opt}} =
\{\vp^* \in \widetilde{\cp}_{\textup{trct}} :  \vp^* \notin [\widetilde{\vp},\boldsymbol{\infty}) \ \forall \widetilde{\vp}\in  \widetilde{\cp}_{\textup{trct}} \setminus \{\vp^*\} \}.
\]
\end{theorem}

>>>>

\begin{definition}
    A problem is defined as $T$-tractable with parameter $\vp$ iff there exists a positive constant $C_{\vp}$, which is independent of $\varepsilon$ and $d$, such that
\begin{equation} \label{eq:tract_def}
	\comp(\varepsilon,d) \le C_{\vp}\, T(\varepsilon^{-1},d,\vp) \qquad \forall \varepsilon >0, \ d \in \mathbb{N}.
\end{equation}
A problem is \emph{strongly}
$T$-tractable with parameter $\vp$ iff the complexity is independent of the dimension of the problem, that is, there exists a positive constant $C_{\vp}$, again independent of $\varepsilon$ and $d$, such that
\begin{equation} \label{eq:strong_tract_def}
	\comp(\varepsilon,d) \le C_{\vp}\, T(\varepsilon^{-1},1,\vp) \qquad \forall \varepsilon >0 , \ d \in \mathbb{N}.
\end{equation}
We allow $\vp$ to be a scalar or vector as the situation dictates.
\end{definition}

@@

If \eqref{eq:tract_def} holds for some $\vp$, it clearly holds for larger $\vp$.  We are often interested in the optimal or smallest $\vp$ for which \eqref{eq:tract_def} holds.  Define the closures of the sets of parameters for which our (strong) tractability conditions hold:
\begin{equation*}
	\cp_{\textup{trct}} : = \{\vp^* : \eqref{eq:tract_def} \text{ holds }\forall \vp \in (\vp^*,\boldsymbol{\infty})\}, \qquad
	\cp_{\textup{strct}} : = \{\vp^* : \eqref{eq:strong_tract_def} \text{ holds }\forall \vp \in (\vp^*,\boldsymbol{\infty})\}.
\end{equation*}

\begin{definition}
    The  set of optimal parameters is defined as all of those parameters satisfying the (strong) tractability conditions that are not greater than or equal to others:
\begin{gather}
	\label{eq:Poptdef}
	\mathcal{P}_{\textup{opt}} : = \{\vp^* \in \cp_{\textup{trct}} :  \vp^* \notin [\vp,\boldsymbol{\infty}) \ \forall \vp \in  \cp_{\textup{trct}} \setminus \{\vp^*\} \}, \\
	\label{eq:Psoptdef}
	\mathcal{P}_{\textup{sopt}} : = \{\vp^* \in \cp_{\textup{strct}} :  \vp^* \notin [\vp,\boldsymbol{\infty}) \ \forall \vp \in  \cp_{\textup{strct}} \setminus \{\vp^*\} \}.
\end{gather}
\end{definition}

@@

Suppose that the domain of interest $(\varepsilon^{-1},d)$ is not all of $(0,\infty) \times \naturals$ but some subset
\begin{equation}
    \Omega  := \{(\varepsilon^{-1},d) : \varepsilon > E(d), \ d \in \naturals \}, \qquad \text{where }E : \naturals \to [0,\infty).
\end{equation}
We then expect that the equivalent conditions for (strong) tractability will be similar in form, but weaker than, the conditions in Theorems  \ref{thm_main_strong_tract2} and \ref{thm_main_tract2}.

\begin{definition}
    A problem is strongly
$T$-tractable with parameter $\vp$ \emph{on the restricted domain $\Omega$} iff the complexity is independent of the dimension of the problem, that is, there exists a positive constant $C_{\vp}$, again independent of $\varepsilon$ and $d$, such that
\begin{equation} \label{eq:rest_strong_tract_def}
	\comp(\varepsilon,d) \le C_{\vp}\, T(\varepsilon^{-1},1,\vp) \qquad \forall(\varepsilon^{-1},d) \in \Omega.
\end{equation}
%\peter{[I am wondering whether this makes sense if $E$ depends on $d$? Do Gnewuch and Wo\'{z}niakowski allow dependence of $E$ on $d$?]}
A problem is 
$T$-tractable with parameter $\vp$ \emph{on the restricted domain $\Omega$} iff the complexity is independent of the dimension of the problem, that is, there exists a positive constant $C_{\vp}$, again independent of $\varepsilon$ and $d$, such that
\begin{equation} \label{eq:rest_tract_def}
	\comp(\varepsilon,d) \le C_{\vp}\, T(\varepsilon^{-1},d,\vp) \qquad \forall(\varepsilon^{-1},d) \in \Omega.
\end{equation}
 Define the closures of the sets of parameters for which our (strong) tractability conditions hold:
\begin{equation*}
	\cp_{\textup{rtrct}} : = \{\vp^* : \eqref{eq:rest_tract_def} \text{ holds }\forall \vp \in (\vp^*,\boldsymbol{\infty})\}, \qquad
	\cp_{\textup{rstrct}} : = \{\vp^* : \eqref{eq:rest_strong_tract_def} \text{ holds }\forall \vp \in (\vp^*,\boldsymbol{\infty})\}.
\end{equation*}
    The  set of optimal parameters is defined as all of those parameters satisfying the (strong) tractability conditions that are not greater than or equal to others:
\begin{gather}
	\label{eq:Poptdef}
	\mathcal{P}_{\textup{ropt}} : = \{\vp^* \in \cp_{\textup{rtrct}} :  \vp^* \notin [\vp,\boldsymbol{\infty}) \ \forall \vp \in  \cp_{\textup{rtrct}} \setminus \{\vp^*\} \}, \\
	\label{eq:Psoptdef}
	\mathcal{P}_{\textup{rsopt}} : = \{\vp^* \in \cp_{\textup{rstrct}} :  \vp^* \notin [\vp,\boldsymbol{\infty}) \ \forall \vp \in  \cp_{\textup{rstrct}} \setminus \{\vp^*\} \}.
\end{gather}
\end{definition}

Define $M: \naturals \to \natzero \cup \{\infty\}$
\begin{equation} \label{eq:Mddef}
    M(d) := \begin{cases}
        \min\{n \in \natzero : \lambda_{n+1,d} \le E(d)\}, & E(d) > 0, \\
        \infty, & E(d) = 0.
    \end{cases}
\end{equation}
This means that 
\[
\left \{
\begin{array}{lll}
&(\lambda_{1,d}^{-1},d), (\lambda_{2,d}^{-1},d), \ldots \notin \Omega, & M(d) = 0, \\
\lambda_{1,d}\ge  \cdots \ge \lambda_{M(d),d} > E(d) \ge\lambda_{M(d)+1,d}, &
(\lambda_{1,d}^{-1},d), \ldots, (\lambda^{-1}_{M(d),d},d)  \in \Omega, & 0 < M(d) < \infty, \\
&
(\lambda_{1,d}^{-1},d), (\lambda_{2,d}^{-1},d), \ldots \in \Omega, & M(d) = \infty.
\end{array}
\right .
\]
Also, for each $d \in \naturals$ define the interval 
\begin{equation} \label{eq:hIdef}
\hI_{d}:=\begin{cases} 
\emptyset, & M(d) = 0, \\
\{1,\ldots, M(d)\}, & 0 < M(d) < \infty, \\
\naturals, & M(d) = \infty.
\end{cases}
\end{equation}
From the definition of $M(d)$ it follows that $(\lambda_{i,d}^{-1},d) \in \Omega  \iff i \in \hI_d$, and
\[
\comp(\varepsilon,d) \in \hI_d \qquad \forall (\varepsilon^{-1},d) \in \Omega.
\]

The equivalent conditions for (strong) $T$-tractability on a restricted domain in Theorems \ref{thm_main_rest_strong_tract} and \ref{thm_main_rest_tract} mimic the conditions given in Theorems \ref{thm_main_strong_tract2} and \ref{thm_main_tract2}, except that the upper limits on the sums now correspond to $M(d)$ rather than $\infty$.  The proofs are a bit more delicate, but similar arguments are used.

We note that it may be possible to define $\Omega$ such that $\lambda_{1,d} > E(d)$ for all $d \in \naturals$. In this case, $M(d) = 0$ for all $d \in \naturals$, the equivalent conditions for (strong) tractability are trivially satisfied, and the zero algorithm satisfies the error tolerance.

\begin{theorem}\label{thm_main_rest_strong_tract}
Let $T$ be a tractability function as specified in \eqref{eq:Tspec} and satisfying \eqref{eq:Tconditions} and \eqref{eq:ptauassume}.  A problem is strongly $T$-tractable \emph{on the restricted domain $\Omega$} iff there exists $\vp \in [\vzero, \boldsymbol{\infty})$ and an integer $L_{\vp} > 0$ such that
\begin{equation} \label{eq:strong_rest_tractiff}
     S_{\vp}:=\sup_{d \in \naturals} \sum_{i = L_{\vp}}^{M(d)} \frac{1}{T(\lambda_{i,d}^{-1},1,\vp)} < \infty.
\end{equation}
By convention, if $M(d) = 0$, the sum is zero.

If \eqref{eq:strong_rest_tractiff} holds for some $\vp$, let  $\widetilde{\cp}_{\textup{rstrct}} : = \{\vp^* : \eqref{eq:strong_rest_tractiff} \text{ holds }\forall \vp \in (\vp^*,\boldsymbol{\infty})\}$.  Then $\cp_{\textup{rstrct}} = \widetilde{\cp}_{\textup{rstrct}}$, and the set of optimal strong tractability parameters on the restricted domain is
\[
	\cp_{\textup{rsopt}} =
	\{\vp^* \in \widetilde{\cp}_{\textup{rstrct}} :  \vp^* \notin [\widetilde{\vp},\boldsymbol{\infty}) \ \forall \widetilde{\vp}\in  \widetilde{\cp}_{\textup{rstrct}} \setminus \{\vp^*\} \}.
\]
\end{theorem}

\begin{proof}
\textbf{Sufficient condition:}\newline
Following the argument for the sufficient condition in Theorem \ref{thm_main_strong_tract2} it follows that for all $d' \in \naturals$ and integers $n$ and $N$ with $1 \le N \le n \le M(d)$

@@

\begin{theorem}\label{thm_main_rest_tract}
Let $T$ be a tractability function as specified in \eqref{eq:Tspec} and satisfying \eqref{eq:Tconditions} and \eqref{eq:ptauassume}.  A problem is $T$-tractable on $\Omega$ iff there exists $\vp\ge \vzero$ and a positive integer $L_{\vp}$ such that
\begin{equation} \label{eq:rest_tractiff4}
     S_{\vp}:=\sup_{d \in \naturals}
     \sum_{i = \lceil L_{\vp}\, T(0,d,\vp) \rceil}^{M(d)} \frac{1}{T(\lambda^{-1}_{i,d},d,\vp)}< \infty.
\end{equation}
If \eqref{eq:rest_tractiff4} holds for some $\vp$, let $\widetilde{\cp}_{\textup{trct}} : = \{\vp^* : \eqref{eq:tractiff4} \text{ holds }\forall \vp \in (\vp^*,\boldsymbol{\infty}) \}$.
Then $\cp_{\textup{trct}} = \widetilde{\cp}_{\textup{trct}}$, and the set of optimal  tractability parameters is
\[
\cp_{\textup{opt}} =
\{\vp^* \in \widetilde{\cp}_{\textup{trct}} :  \vp^* \notin [\widetilde{\vp},\boldsymbol{\infty}) \ \forall \widetilde{\vp}\in  \widetilde{\cp}_{\textup{trct}} \setminus \{\vp^*\} \}.
\]
\end{theorem}
